---
title: "rXTalkViz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rXTalkViz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction
The rXTalkViz package provides tools for quantifying and visualizing cross-talks between biological pathways by employing. It models the pathway cross-talks in two categories, 1) via share nodes (genes/proteins/molecules) - defined as Type-I cross, and 2) via shared interactions (Protein-Protein interaction, or data-driven interactoins) - defined as Type-II cross-talks. Here, the Type-I and Type-II cross-talks are quantified with Jaccard-Index score, and Network proximity scores, respectively. Visualization includes Network plots, Shankey diagram, and Heatmap plots. Finally, the whole quantification result is also saved as spreadsheet (as a csv file). This vignette will guide you through the basic usage of the package, including how to load the data, perform cross-talk analysis, and visualize the results.

## Step 1: Loading the Required Libraries
First, we need to load the necessary libraries, including rXTalkViz and other dependencies such as dplyr and data.table.

```{r setup, message=FALSE, warning=F}
library(rXTalkViz)
library(dplyr)
library(data.table)
```

## Step 2: Loading Example Data
The package comes with an example dataset named example_enrich.df. Let's observe the head of this data to understand its structure.

```{r, message=FALSE, warning=F}
print(head(example_enrich.df))
```

This dataset contains enrichment results for various pathways. We will use this data to explore the cross-talk analysis features of the package.

## Step 3: Filtering Data
To focus on the most significant pathways, we will filter the dataset to include only those pathways with an adjusted p-value (p.adjust) below 0.001. Moreover, this will greatly control the time required for running the whole pipeline by invoking the wrapper function.


```{r, message=FALSE, warning=F}
filtered_data <- example_enrich.df %>%
  filter(p.adjust < 0.001)
filtered_data %>% nrow
```

## Step 4: Performing Cross-Talk Analysis
With the filtered data, we can now perform cross-talk analysis using the xTalk_wrapper function. This function allows you to quantify, visualize cross-talks between pathways, and save all the results within the specified directory. Note, if the directory doesn't exist, it will create on but with a warning, and the directory name can be specified by changing the 'outdir' parameter.

```{r out.width="100%", out.height="100%", message=FALSE, warning=F}
# filtered_df <- example_enrich.df %>%
#   filter(p.adjust < 0.001)
# xTalk_wrapper(filtered_df,
#               doPlot = T,
#               doXtalkQuant = T,
#               nPermute = 2,
#               min_cross_talk_score = 1.0,
#               plot_width = 10,
#               plot_height = 10)
```

```{r fig.width=10, fig.height=10, message=FALSE, warning=F}
filtered_df <- example_enrich.df %>%
  filter(p.adjust < 0.001)

p <- xTalkPlot.Type_I.NetworkView(
  enrich.df = filtered_df,
  string_PPI_score_th = string_PPI_score_th,
  showCategory = 100,
  layout = "linear",
  colorEdge = TRUE,
  circular = TRUE)
p
```


In this network plot, Type-I cross-talks is shown among highly enriched pathways via shared genes.

# Conclusion
In this vignette, we demonstrated the basic functionality of the rXTalkViz package. We loaded example data, filtered it based on significance, and performed cross-talk analysis. For more detailed analysis, refer to the package documentation.
